---
title: "Micrarray Analysis for Gene Expression from a Local **series_matrix** File (GSE26471, 1 samples)"
author: "Florent Chuffart"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Loading data using Local series_matrix File}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Loading the epimedtools library

``` {r}
library(epimedtools)
```

Creating a new study that will be cached in the file /tmp/study.rds
```{r}
study = create_study("/tmp/study.rds")
```

This study deals with the  GEO accession GSE26471, its series_matrix is distributed with the package
``` {r}
study$series_matrix_filename = system.file(
  "extdata/GSE26471", 
  "GSE26471_series_matrix.txt.gz", 
  package = "epimedtools"
)
```

Load data associated with this study. Data is even on the hard drive, it is directly load from there.
``` {r}
study$get_gset()
```

Extract data matrix.
``` {r}
dim(study$get_data())
```

Extract experimental grouping dataframe.
``` {r}
dim(study$get_exp_grp())
```

Plot density of expression for this *1-*sample dataset. 
``` {r, fig.width=7, fig.height=5}
plot(density(study$get_data()[,1]))
```
