---
title: "Mann-Whitney Gene Expression analysis"
author: "Florent Chuffart"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mann-Whitney Gene Expression analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Loading the epimedtools library

```{r}
library(epimedtools)
```

Creating a new study.

```{r}
study = create_study()
```

The .CEL.gz files are stored into two directories the *ctrl* directory and the *kc* directory. When processing .CEL.gz files, the origin of each file  (one of this two directories) is stored in a new column of the experimental grouping. This column is next used used to perform the Mann-Witney test, one `orig` vs. the otehr.  

```{r}
kc_cel_filedir = system.file(
  "extdata/trscr_raw_kc", 
  package = "epimeddata"
)
print(kc_cel_filedir)
ctrl_cel_filedir = system.file(
  "extdata/trscr_raw_ctrl", 
  package = "epimeddata"
)
print(ctrl_cel_filedir)
study$cel_filedirs = c(kc_cel_filedir, ctrl_cel_filedir)
print(study$cel_filedirs)
```

The RMA normalized data matrix is extract using the **get_data** method.

```{r}
head(study$get_data())
```

The associated clinical data

```{r}
print(study$get_exp_grp())
```

The Mann-Whitney analysis on 100 first probes.

```{r}
# selecting intersting probes
probe_names = rownames(study$get_data())[1:100]
# subset dataset
ctrl_exp_grp_key = "orig"
case_exp_grp_key = "orig"
ctrl_factor_name = "trscr_raw_ctrl"
case_factor_name = "trscr_raw_kc"

mw_res = study$do_mw_test(probe_names, ctrl_exp_grp_key, case_exp_grp_key, ctrl_factor_name, case_factor_name)

thres = c(1,-1) * -log10(0.05)
fc_thres = c(-1.5, -1.2, 1.2, 1.5)
lr_thres = gtools::foldchange2logratio(fc_thres)

# Whitout FDR
plot(mw_res$trscr_raw_ctrl.trscr_raw_kc.logratio, -log10(mw_res$trscr_raw_ctrl.trscr_raw_kc.mw_pval_adj))


# Whit FDR 
# plot(mw_res$mean_fc, -log10(p.adjust(mw_res$mw_pval, "BH")))
abline(h=-log10(0.05), v = lr_thres, lty=3, col=2)
legend("bottomright", lty=3, col=2, legend=paste("fc thres (", paste(fc_thres, collapse=","), ")", sep=""))

```
