---
title: "Micrarray Analysis for Gene Expression from Local .CEL.gz Files"
author: "Florent Chuffart"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Loading data using local .CEL.gz files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Loading the epimedtools library

``` {r}
library(epimedtools)
```

Creating a new study.
```{r}
study = create_study(Study_RC_name="Study_raw_trscr")
```

The .CEL.gz files are stored into two directories the *ctrl* directory and the *kc* directory:
``` {r}
kc_cel_filedir = "../inst/extdata/trscr_raw_kc"
ctrl_cel_filedir = "../inst/extdata/trscr_raw_ctrl"
study$cel_filedirs = c(kc_cel_filedir, ctrl_cel_filedir)
print(study$cel_filedirs)
```

The experimental grouping of **kc** samples is laoded before loading data.
``` {r}
exp_grp = read.table(file=gzfile("../inst/extdata/trscr_raw_kc/expgrp_kc.csv.gz"), stringsAsFactors=FALSE, sep=";", header=TRUE)
rownames(exp_grp) = paste(exp_grp$sample, ".CEL.gz", sep="")
study$exp_grp = exp_grp
```

The RMA normalized data matrix is extract using the **get_data** method.
``` {r}
head(study$get_data())
```

The origin of sample is injected into the experimental grouping thank to the column **orig**.
``` {r}
study$get_exp_grp()
```

This study deals with data obtained using the **GPL570** platform.
``` {r}
study$platform_name = "GPL570"
```

Plot density of expression for this *1-*sample dataset.
``` {r, fig.width=7, fig.height=5}
plot(density(study$get_data()[,1]))
```

Plot density of expression for this *1-*sample dataset using ratio.
``` {r, fig.width=7, fig.height=5}
ctrl_sample_names = rownames(study$exp_grp)[study$get_exp_grp()$orig == "trscr_raw_ctrl"]
plot(density(study$get_ratio(ctrl_sample_names)[,1]))
```

Saving the study on the file system.
```{r}
study$save("/tmp/study.rds")
```
