---
title: "Micrarray Analysis for Gene Expression from a Local Tabuled Files"
author: "Florent Chuffart"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Loading data using local tabuled files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Loading the epimedtools library

``` {r}
library(epimedtools)
```

Creating a new study.
```{r}
study = create_study(Study_RC_name="Study_loc")
```

This study deals with data obtained using the **GPL570** platform.
``` {r}
study$platform_name = "GPL570"
```

The experimental grouping is stored into a compressed text file distributed with the package:
``` {r}
exp_grp_filename = system.file(
  "extdata/GSE26471", 
  "GSE26471_exp_grp.tab.gz", 
  package = "epimedtools"
)
print(exp_grp_filename)
study$exp_grp = read.table(file=gzfile(exp_grp_filename), stringsAsFactors=FALSE)
```


The expression data matrix is stored into a compressed text file distributed with the package:
``` {r}
data_filename = system.file(
  "extdata/GSE26471", 
  "GSE26471_data.tab.gz", 
  package = "epimedtools"
)
print(data_filename)
study$data = as.matrix(read.table(file=gzfile(data_filename)))
```

Extract data matrix.
``` {r}
dim(study$get_data())
```

Extract experimental grouping dataframe.
``` {r}
dim(study$get_exp_grp())
```

Plot density of expression for this *1-*sample dataset. 
``` {r, fig.width=7, fig.height=5}
plot(density(study$get_data()[,1]))
```

Saving the study on the file system.
```{r}
study$save("/tmp/study.rds")
````
